Learning Actionable Analytics (with Applications for Reducing Defects and Reducing Runtimes)

### 2015-05-01 manuscript.tex
Updates from ShareLaTeX

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath} 
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=#1,
numberstyle=\footnotesize,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\small\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, WHAT, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
%
}}{}

\begin{document}
  \title{Temporal Transfer Learning with ``HERE'':\\ Local Policy Generation (in Near Linear Time)}

@@

  \begin{abstract}
 Busy project managers grow impatient with analytcs
  \end{abstract}
  \begin{IEEEkeywords}
    defect prediction, CART
  \end{IEEEkeywords}

\section{Introduction}

In the $21^{st}$ century, it is now impossible to manually browse all
the available software defect data.  For example, as of April 2015,
Mozilla Firefox has over a million bug reports on software projects.  While it is
now possible to automatically analyze such data with data miners, at
some stage a group of business users will have to convene to {\em
  interpret the results} (e.g., to decide if it is wise to deploy the
results as a defect reduction method within an organization).
These business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining.

In order to satisfy these panelists, w  explore the generation of {\em policies} in the context
of software analytics. Such {\em policies} are recommendations of what
to change in order to better achieve some goal. In this paper the two
case studies are:
\bi
\item Learning code changes to static code features that reduce
the number of defects;
\item Learning reconfigurations of Makefiles such that, when the code
is compiled, that code runs faster.
\ei
Note that {\em policy generation} is different to the standard
data mining tasks of prediction or classification.  Predictors
and classifiers only tell us what {\em is} while policy 
generators tell us {\em what to do}.

The tool used to generate these policies is ``HERE'', an instance-based
local search algorithm that looks for improvement trends in the data,
then adjusting examples to nudge them along that improvement. While
our case studies are defect and runtime reduction, HERE is a general
purpose algorithm that could be applied to any table of data with at least
one column whose values can be sorted from ``desired'' to ``undesired''.

@@

We also distinguish HERE's policy generation fro the multi-objective optimization methods explored by Harman and Krall and Menzies and 
and many ithers.

@@

{\bf RQ2: Can lessons learnt from local search operators be useful for generating solutions that can be used to mitigate defects in software systems?}

@@

An alternative approach would be to make use of an instance based approach, an example of case based reasoning strategy, in place of the conventional model based approach. Instance based approaches have been used extensively by the effort estimation community, see~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}. This approach has been proposed as an alternative to closed form mathematical models or other modeling methods such as regression \cite{keung2008analogy}. There are several other reasons for instance based approaches being a useful tool, see~\cite{6600685}. As pointed out by~\cite{walkerden1999empirical} it can be used with partial knowledge of the target project at an early stage which could be a very useful tool in preventing software defects. Instance based approaches are also rather robust in handling cases with sparse samples \cite{1438374}. All these features are desirable and suggest that instance-based approach is a useful adjunct to traditional model based approach.

@@

\subsection{Planning changes using WHAT}
We propose the WHAT algorithm as a tool to plan changes in the original test data. Figure~\ref{fig:what} highlights the procedure by which WHAT operates. It begins by creating clusters which are generated using the WHERE algorithm discussed above. Following this, a nearest neighbour scheme is applied to identify pairs of nearby clusters. A projective plane is constructed with these clusters at the vertices to characterize the cluster pairs.

The mutation policy works by projecting the test instance onto the projective planes and identifying the the plane on which the test instance has the largest scalar projection. The planner then reflects over the vertices of the chosen hyperplane, identifying the \textit{better} vertex among the two. Now, a new instance is generated by \textit{mutating the attributes} of the test instance towards the better vertex and away from the worse vertex. This process is repeated for all the test instances that are considered defective by the defect prediction scheme, discussed below.

\subsection{Feature Weighting}
The algorithm also allows for a feature weighting and pruning scheme. We have employed a simple (and fast) attribute ranking method where the features are sorted based on their ability to find meaningful splits in the dependent class of the data, see ~\cite{hall03}. Features that can find splits with the \textit{lesser variance} in the dependent variable are \textit{ranked higher} than features that don't.

We reason that the most informative feature must undergo a larger change as compared to other features. Therefore, when feature weighting is employed, the higher ranked features are mutated by a larger extent. Additionally, we have also included information pruning. Here we mutate only the top X\% of the features in each test instance. The remaining features remain unaltered. 
\begin{figure}[htbp!]
\small
\begin{tabular}{|p{.95\linewidth}|}\hline
WHAT begins by \textit{clustering} the training data into similar subsets using WHERE. WHERE works as follows:
\begin{enumerate}
\item Find the two most dist points in a given population; say {\em east} and {\em west}. 
\item Draw an axis of length $c$ between the poles. 
\item Let each point be at distance $a,b$ to the {\em east,west} poles.  Using the cosine rule, project each point onto the  axis  at $x=(a^2 + c^2 - b^2)/(2c)$.  
\item Using the Medianian $x$ value, divide the population.
\item For each half that is larger than $\sqrt{N}$ of the original population, go to step 1.
\end{enumerate}

@@

The final clusters that are found by WHERE are then used by WHAT as follows:
\begin{itemize}
\item For each cluster find its nearest neighbor.
\item Construct a projective plane for each cluster pair. Label the cluster pairs as {\em Good} and {\em Bad}.
\item For a test instance, project it onto the planes generated above and pick the one on which the test case has the largest projection.
\item Mutate the attributes of the test case towards the {\em Good Cluster} in that projective plane. 
\end{itemize}

Formally, WHAT can be considered as an instance based  learning scheme. And that's because it builds a set of recommendations based on a given test case based on the clusters for Median at the training stage. This is unlike a model based approach, which would have used a model to do the same~\cite{xomo}.
\\\hline
\end{tabular}
\caption{Inside WHAT}\label{fig:what}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{_figs/WHAT-Clusters2.pdf}
\caption{WHAT}
\label{fig:whatflow}
\end{figure}

@@

def project(west, east, c, x): 
  "Project x onto line east to west"
  a = dist(x,west)
  b = dist(x,east)
  #cosine rule
  return (a*a + c*c - b*b)/(2*c)

@@

def getContours(train):
  """
  Identify the pairs of clusters
  that can form a contour
  """
  contours = []
  # Clustering with where
  clusters = WHERE(train) 
  near = lambda x: knn(x, clusters)[1]
  for c in clusters:
    one = envied(c)
    two = envied(near(one, clusters))
    contours += [one, two]
  return contours

def project(a, b, test):
  if score(a) > score(b):
    good, bad = a,b 
  else:
    good, bad = b,a
  AB = dist(bad,test)
  BC = dist(good,test)
  AC = dist(good,bad)
  # Cosine rule
  y = (AB$^2$ + BC$^2$ - AC$^2$)/(2*AC)
  return $\sqrt{AB^2 - y^2}$

@@

def WHAT():
  contours = getContours()
  weights = fWeight(train)
  for rows in defectiveRows:
    vtx = nearestContour(contours, rows)
    [good, bad] = sorted(vtx
            , key=lambda F: score(F))
    new += mutate(rows, good, weights)
  return new
\end{python}
\caption{Instance-based Planning using WHAT}
\label{fig:WHAT}   
\end{figure}

@@

In short, SMOTE works by under-sampling the majority class and oversampling all the minority classes in the training data. We use a similar approach with one minor addition to the original algorithm. In our implementation of SMOTE we have introduced an additional step called \textit{resampling}, wherein we ensure that after we do the over/under sampling, the new training data does not have any of the original rows. The new training data merely resembles the original data. This is done as a precautionary measure, so as not to have WHAT and Random Forest train on the same training data. 
%Its worth noting that both WHAT and Random Forest require some data to train. In order to prevent the two from using the same data to train, we run SMOTE by  \textit{resampling} the classes. Re-sampling

@@

\caption{Attributes of the Performance Prediction datasets}\label{fig:cpm}
\end{figure}

@@

%   \subfloat[Experimental Rig for Defect Prediction]{  \includegraphics[width=\linewidth]{_figs/WHAT}
%   \label{fig:rig2}}

@@

The following section describes the experimental rig and the experiments used to measure the performance of WHAT on 10 defect data sets and 6 performance prediction data sets.

\subsection{Defect Data Set}
The data was obtained from the PROMISE repository\footnote{Promise Repository: \url{http://openscience.us/repo}}. For the defect data we investigated 32 releases from 11 open source Java projects defined by the metrics highlighted in figure ~\ref{fig:ck}: \kw{Apache } (1.5 -- 1.7), \kw{Apache Camel} (1.2 -- 1.6), \kw{Apache Ivy} (1.1 -- 2.0), \kw{JEdit} (4.1 -- 4.3), \kw{Apache Log4j} (1.0 -- 1.2), \kw{Apache Lucene} (2.0 -- 2.2), \kw{PBeans} (1.0 and 2.0), \kw{Apache POI} (2.0 -- 3.0), \kw{Apache Synapse} (1.0 -- 1.2), \kw{Apache Velocity} (1.4 -- 1.6), and \kw{Apache Xalan-Java} (2.5 -- 2.7).

@@

The experimental rig for defect prediction is shown in Figure \ref{fig:rig}. It works as follows: 
\bi
\item We use an {oracle} (Random Forest + SMOTE) to determine whether a certain test case is defective or not. 
\item If the oracle suggests that an instance is defective, then we use WHAT to apply the recommendations to that test instance. 
\item The oracle is then used to predict the attributes in the modified instance.
\item The statistical significance of the changes are verified using the cliffs delta measure described below.
\ei

For run time optimization, we use a similar rig. The oracle is Random Forest, and it predicts for run times of specific configuration. WHAT is used to generate a recommended configuration to reduce the run times. The run time is measured by the oracle. The reduction of run times are expressed as fractions of the original (also referred to as baseline) run time.

@@

In order to assess the performance of the planner for the defect data set, we use the Cliffs Delta score to measure the probability that the number of bugs in test data before applying the planner is larger than after doing so. In other words, we use the delta score to measure the ability of the planner to effective reduce the number of defects. Given the untreated test instance labeled \textit{Before} of length \textit{M} and the treated instances labeled \textit{After}, the delta score is obtained as follows:
\begin{equation}
\delta = \frac{\#(Before<After) - \#(Before>After)}{M^2}
\end{equation}

@@

In addition to the above, we rank the different varis of the planning scheme to identify the best approach. We make use of the Scott-Knott procedure, recommended by Mittas \& Angelis in their 2013 IEEE TSE paper~\cite{sk}, to compute the ranks. It works as follows: A list of treatments \textit{l} is sorted by the Median score. The list \textit{l} is then split into sub-lists m, n in order to maximize the expected value of the differences in the observed performance before and after division. A statistical hypothesis test \textit{H} is applied on the splits \textit{m, n} to check if they are statistically different. If so, Skott-Knott then recurses on each division.

@@

\section{Experimental Results}
\begin{figure}[!t]
\centering
\includegraphics[width=\linewidth]{_figs/histograms.pdf}
\caption{Comparison of defects before and after planning.}
\label{fig:bugs}
\end{figure}

>>>>

 \documentclass[conference]{IEEEtran}
\usepackage{subfig}
\usepackage{wrapfig}
 \usepackage{amsmath}
 \usepackage{url}
 \usepackage{pifont}
 \usepackage{times}
\usepackage{rotating}
%\usepackage{balance} 
\usepackage{color, colortbl}
\usepackage{graphicx}
\usepackage{algorithmicx}
\usepackage{program}
\usepackage{cite}
\usepackage{alltt}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sec:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\definecolor{lightgray}{gray}{0.975}
\usepackage{fancyvrb}
\usepackage{stfloats}
\usepackage{multirow}
\usepackage{listings}
\usepackage{amsmath} 
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}

@@

\lstnewenvironment{python}[1][]{
\lstset{
mathescape,
numbers=#1,
numberstyle=\footnotesize,
numbersep=0.5em,
xleftmargin=0em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\small\setstretch{0.8},
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap, project, furthest, split, WHERE,clusterer, getContours, envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
%
}}{}

\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
Rahul Krishna, Tim Menzies, Xipeng Shen\\
       \affaddr{Computer Science, NcState, USA}\\
       \{i.m.ralk, tim.menzies, xipengshen\}@gmail.com
% 3rd. author
 % use '\and' if you need 'another row' of author names
% 4th. author
\and
 Andrian Marcus\\
       \affaddr{Computer Science, UtDallas, USA}\\
       {amarcus@utdallas.edu} }

\begin{document}
  \title{Learning Actionable Analytics 
 (with Applications for Reducing Defects and Reducing Runtimes)}

@@

  \begin{abstract}
 As business users demand more insightful
 analytics, data scientists need to change
 their tools. Instead of merely predicting 
 some result, they also need tools that generate ``policy advice'';
 i.e. suggestions on  how to change values in order to
 improve on some predicted value.

 This paper proposes one such policy
 generator. ``HOW'' is a near-linear-time
 tool for proposing changes to independent
 variables in order to improve on 
 the predictions of the dependent variables. Unlike other approaches
 for learning optimizations to software does not require
 an underlying of the domain (rather, HOW automatically
 learns that model as part of its operation).

 This paper uses  HOW to learn methods
 to reduce defects and decrease program runtime.
 For the test data used in this paper,
 we have observed that the improvements generated by HERE can reduce
 defect prediction and runtimes by up to
 50\% of their original values.
  \end{abstract}
  \begin{IEEEkeywords}
Instance-based reasoning, defect prediction, random forest, optimization.
  \end{IEEEkeywords}

\section{Introduction}
In the $21^{st}$ century, it is now impossible to manually browse all
the available software defect data.  For example, as of April 2015,
Mozilla Firefox has over a million bug reports on software projects.  While it is
now possible to automatically analyze such data with data miners, at
some stage a group of business users will have to convene to {\em
  interpret the results} (e.g., to decide if it is wise to deploy the
results as a defect reduction method within an organization).

These business  users are now demanding that data mining tools
be augmented with tools to support  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
that follow the mining, rather than just  the mining. For example:
\bi
\item Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users now demand {\em policies} that
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
{\em policies} that reduce that runtime.
\ei
In order to satisfy the demands of this new kind of business user, we  explore the generation of {\em policies} in the context
of software analytics. Such {\em policies} are recommendations of what
to change in order to better achieve some goal. 
Note that such {\em policy generation} is very different to the standard
data mining tasks of prediction some discrete or numeric class.
Predictors
and classifiers only tell us what {\em is} while policy 
generators tell us {\em what to change}. For the kind of business user
discussed above, policy generators are the prefeered tool since,
as one gruff user once shouted at us ``don't tell me what \underline{is};
tell me what to \underline{change}''.

This paper presents the following general framework for extending  predictor systems
such that they become 
policy generators:
\begin{small}
\[
project\;data \rightarrow \left\{\begin{array}{l}\mathit{train}
\rightarrow \left\{\begin{array}{l}\mathit{predictor\;(e.g.\;CART\;or\; RF)}\\
\mathit{policy\; generator\; (e.g.\; HOW)}\end{array}\right.
\\
~\\
\mathit{test}  \rightarrow  \left\{\begin{array}{r@{~}l}
\mathit{before}& =\mathit{predictor}(\mathit{test})\\
\mathit{after}& =\mathit{predictor}(\; \mathit{policy}(\mathit{test})\;)\\
\Delta& =\mathit{after} / \mathit{before}
\end{array}\right.
\end{array} \right.
\]
\end{small}
In the above:
\bi
\item
Project data is divided into a train and test set.
\item
From the train set, we learn {\em both} a policy generator as well
as a predictor using (e.g.)
CART (for making numeric predictions~\cite{Breiman1984}) or Random Forests (for making predictions
about symbolic classes~\cite{Breiman2001}). 
\item
The test data is then studied to generate two predictions:
a {\em before} prediction (on the raw test data); then an {\em after}
prediction (after applying the policies learned from the training data).
\item
The improvement delta $\Delta$ is then computed by dividing {\em after} and {\em before}.
\ei
This paper presents a novel policy generator called ``HOW''.
HOW is a near-linear time  instance-based
local search algorithm that looks for improvement trends in the data,
then adjusting examples to nudge them along that trend.
The experiments of this paper apply HOW to learning policies for 
software defect reduction and for compiling software that runs faster.
At least for the data explored here, HOW's reductions in defects and runtimes
are very large-- up to 50\% fewer defects and up to 50\% faster runtimes.

The rest of this paper is structured as follows.

\section{Releated Work}

Our prior work in generating policies relied upon a Monte Carlo
analysis of an underlying model~\cite{me07f}. This approach had three major problems.
Firstly, enough domain knowledge had to be collected to build that model
and this can take a long time. For example, in our previous work we used three software process models from
the Unviersity of Southern California that were initially
developed and updated for nearly two decades from 1981~\cite{boehm81} to 2000~\cite{boehm00b}.

Secondly, there is the computational cost of running those models.
While

~\cite{}-- which suffered from the problem of (a)~the effort required to collect the domain knowledge to build
problem in certifying the underlying model;
 (b)~the computational problem of running that model enough
times to learn patterns in that data; and (c)~ requrearea used large scale simulations or large

While
our case studies are defect and runtime reduction, HERE is a general
purpose algorithm that could be applied to any table of data with at least
one column whose values can be sorted from ``desired'' to ``undesired''.

@@

We also distinguish HERE's policy generation fro the multi-objective optimization methods explored by Harman and Krall and Menzies and 
and many others.

@@

{\bf RQ2: Can lessons learnt from local policy generators be useful for generating solutions that can be used to mitigate defects in software systems?}

@@

An alternative approach would be to make use of an instance based approach, an example of case based reasoning strategy, in place of the conventional model based approach. Instance based approaches have been used extensively by the effort estimation community, see~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}.

This approach has been proposed as an alternative to closed form mathematical models or other modeling methods such as regression \cite{keung2008analogy}. There are several other reasons for instance based approaches being a useful tool, see~\cite{6600685}. As pointed out by~\cite{walkerden1999empirical} it can be used with partial knowledge of the target project at an early stage which could be a very useful tool in preventing software defects. Instance based approaches are also rather robust in handling cases with sparse samples \cite{1438374}. All these features are desirable and suggest that instance-based approach is a useful adjunct to traditional model based approach.

@@

\subsection{Planning changes using HERE}
We propose the HERE algorithm as a tool to plan changes in the original test data. Figure~\ref{fig:what} highlights the procedure by which HERE operates. It begins by creating clusters which are generated using the WHERE algorithm discussed above. Following this, a nearest neighbour scheme is applied to identify pairs of nearby clusters. A contour is constructed with these clusters at the vertices to characterize the cluster pairs.

The mutation policy works by projecting the test instance onto the contours and identifying the the contour on which the test instance has the largest scalar projection. The planner then reflects over the vertices of the chosen contour, identifying the \textit{better} vertex among the two. Now, a new instance is generated by \textit{mutating the attributes} of the test instance towards the better vertex and away from the worse vertex. This process is repeated for all the test instances that are considered defective by the defect prediction scheme, discussed below.

\subsection{Feature Weighting} \label{fwt}
The algorithm also allows for a feature weighting and pruning scheme. We have employed a simple (and fast) attribute ranking method where the features are sorted based on their ability to find meaningful splits in the dependent class of the data, see ~\cite{hall03}. Features that can find splits with the \textit{lesser variance} in the dependent variable are \textit{ranked higher} than features that don't.

We reason that the most informative feature must undergo a larger change as compared to other features. Therefore, when feature weighting is employed, the higher ranked features are mutated by a larger extent. Additionally, we have also included information pruning. Here we mutate only the top X\% of the features in each test instance. The remaining features remain unaltered. 
\begin{figure}[htbp!]
\small
\begin{tabular}{|p{.95\linewidth}|}\hline
HERE begins by \textit{clustering} the training data into similar subsets using WHERE. WHERE works as follows:
\begin{enumerate}
\item Find the two most dist points in a given population; say {\em east} and {\em west}. 
\item Draw an axis of length $c$ between the poles. 
\item Let each point be at distance $a,b$ to the {\em east,west} poles.  Using the cosine rule, project each point onto the  axis  at $x=(a^2 + c^2 - b^2)/(2c)$.  
\item Using the Median $x$ value, divide the population.
\item For each half that is larger than $\sqrt{N}$ of the original population, go to step 1.
\end{enumerate}

@@

The final clusters that are found by WHERE are then used by HERE as follows:
\begin{itemize}
\item For each cluster find its nearest neighbor.
\item Construct a projective plane for each cluster pair. Label the cluster pairs as {\em Good} and {\em Bad}.
\item For a test instance, project it onto the planes generated above and pick the one on which the test case has the largest projection.
\item Mutate the attributes of the test case towards the {\em Good Cluster} in that projective plane. 
\end{itemize}

Formally, HERE can be considered as an instance based  learning scheme. And that's because it builds a set of recommendations based on a given test case based on the clusters for Median at the training stage. This is unlike a model based approach, which would have used a model to do the same~\cite{xomo}.
\\\hline
\end{tabular}
\caption{Inside HERE}\label{fig:what}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\linewidth]{_figs/WHAT-Clusters2.pdf}
\caption{HERE}
\label{fig:whatflow}
\end{figure}

@@

def project(west, east, AC, x): 
  "Project x onto line east to west"
  AB = dist(x,west)
  BC = dist(x,east)
  #cosine rule
  return (AB$^2$ + AC$^2$ - AB$^2$)/(2*AC)

@@

def getContours(train):
  """
  Identify the pairs of clusters
  that can form a contour
  """
  contours = []
  # Clustering with where
  clusters = WHERE(train) 
  near = lambda x: knn(x, clusters)[1]
  for c in clusters:
    one = envied(c)
    two = envied(near(one, clusters))
    contours += [(one, two)]
  return contours

def project(a, b, test):
  if score(a) > score(b):
    good, bad = a,b 
  else:
    good, bad = b,a
  AB = dist(bad,test)
  BC = dist(good,test)
  AC = dist(good,bad)
  # Cosine rule
  x = (AB$^2$ + AC$^2$ - BC$^2$)/(2*AC)
  return $\sqrt{AB^2 - x^2}$

@@

def HERE():
  contours = getContours()
  weights = fWeight(train)
  for rows in defectiveRows:
    vtx = nearestContour(contours, rows)
    [good, bad] = sorted(vtx
            , key=lambda F: score(F))
    new += mutate(rows, good, weights)
  return new
\end{python}
\caption{Instance-based Planning using HERE}
\label{fig:HERE}   
\end{figure}

@@

In short, SMOTE works by under-sampling the majority class and oversampling all the minority classes in the training data. We use a similar approach with one minor addition to the original algorithm. In our implementation of SMOTE we have introduced an additional step called \textit{resampling}, wherein we ensure that after we do the over/under sampling, the new training data does not have any of the original rows. The new training data merely resembles the original data. This is done as a precautionary measure, so as not to have HERE and Random Forest train on the same training data. 
%Its worth noting that both HERE and Random Forest require some data to train. In order to prevent the two from using the same data to train, we run SMOTE by  \textit{resampling} the classes. Re-sampling

@@

\caption{Attributes of the Performance Prediction data sets}\label{fig:cpm}
\end{figure}

@@

%   \subfloat[Experimental Rig for Defect Prediction]{  \includegraphics[width=\linewidth]{_figs/HERE}
%   \label{fig:rig2}}

@@

The following section describes the experimental rig and the experiments used to measure the performance of HERE on 10 defect data sets and 6 performance prediction data sets.

\subsection{Defect Data Set}
The data was obtained from the PROMISE repository\footnote{Promise Repository: \url{http://openscience.us/repo}}. For the defect data we investigated 32 releases from 11 open source Java projects defined by the metrics highlighted in figure ~\ref{fig:ck}: \kw{Apache Ant} (1.5 -- 1.7), \kw{Apache Camel} (1.2 -- 1.6), \kw{Apache Ivy} (1.1 -- 2.0), \kw{JEdit} (4.1 -- 4.3), \kw{Apache Log4j} (1.0 -- 1.2), \kw{Apache Lucene} (2.0 -- 2.2), \kw{PBeans} (1.0 and 2.0), \kw{Apache POI} (2.0 -- 3.0), \kw{Apache Synapse} (1.0 -- 1.2), \kw{Apache Velocity} (1.4 -- 1.6), and \kw{Apache Xalan-Java} (2.5 -- 2.7).

@@

The experimental rig for defect prediction is shown in Figure \ref{fig:rig}. It works as follows: 
\bi
\item We use an {oracle} (Random Forest + SMOTE) to determine whether a certain test case is defective or not. 
\item If the oracle suggests that an instance is defective, then we use HERE to apply the recommendations to that test instance. 
\item The oracle is then used to predict the attributes in the modified instance.
\item The statistical significance of the changes are verified using the cliffs delta measure described below.
\ei

For run time optimization, we use a similar rig. The oracle is Random Forest, and it predicts for run times of specific configuration. HERE is used to generate a recommended configuration to reduce the run times. The run time is measured by the oracle. The reduction of run times are expressed as fractions of the original (also referred to as baseline) run time.

@@

In order to assess the performance of the planner for the defect data set, we use the Cliffs Delta score to measure the probability that the number of bugs in test data before applying the planner is larger than after doing so. In other words, we use the delta score to measure the ability of the planner to effective reduce the number of defects. Given the untreated test instance labeled \textit{Before} of length \textit{M} and the treated instances labeled \textit{After}, the delta score is obtained as follows:
\begin{equation}
\delta = \frac{\#(Before<After) - \#(Before>After)}{M^2}
\label{eq:cliffs}
\end{equation}

@@

In addition to the above, we rank the different various of the planning scheme to identify the best approach. We make use of the Scott-Knott procedure, recommended by Mittas \& Angelis in their 2013 IEEE TSE paper~\cite{sk}, to compute the ranks. It works as follows: A list of treatments \textit{l} is sorted by the Median score. The list \textit{l} is then split into sub-lists m, n in order to maximize the expected value of the differences in the observed performance before and after division. A statistical hypothesis test \textit{H} is applied on the splits \textit{m, n} to check if they are statistically different. If so, Skott-Knott then recurses on each division.

@@

\section{Experimental Results}

\subsection{Defect Data Set}
This sections explores \textbf{RQ2: Can lessons learnt from local policy generators be useful for generating solutions that can be used to mitigate defects in software systems?}

Figure \ref{fig:bugs} compares the number of defects before and after applying "HERE". The graphs represent the median scores, with the inter-quartile ranges represented by the error bars. The results seem to show that in all data sets under consideration, the number of defects have reduced, some more so than others. Large differences were notice in certain data sets namely, \textit{ant}, \textit{camel}, \textit{jedit}, \textit{poi,} and \textit{xalan}, on close examination it was noticed that these data sets had comparatively large training data, which translates to more samples for the planner to learn from.

In addition to this, to study the statistical significance of the changes, the results were statistically verified using the Skott-Knott procedure. In order to do this we computed the Cliff's Delta score. As highlighted previously, HERE is applied only for test instances that are considered defective by the oracle. Using \eq{cliffs} we can expect to see the following results:

\bi
\item $\lt$ 0 if defects have been suppressed.
\item = 0 if no change has been made.
\item $\gt$ 0 if new defects have been introduced. This is not applicable to our case because we only apply HERE to defective cases; either the defects are suppressed or they remain unaltered, introduction of new defects is not possible.
\ei

\begin{figure}[tbp]
\centering
\includegraphics[width=\linewidth]{_figs/histograms.pdf}
\caption{Comparison of defects before and after planning.}
\label{fig:bugs}
\end{figure}
With this in mind, it is easy to note that lower cliffs delta score translates to better the performance.

The results are tabulated in figure \ref{tab:ant}. For the sake of completeness, we have introduced three basic configurations in HERE: 
\bi
\item \textit{Extent of Mutation:} Denotes the extent by which a test instant is \textit{mutated} towards the best cluster in the contour. It is a fraction that lies between 0 and 1. The baseline study requires no change be made, and thus extent of mutation is set to zero.
\item \textit{Feature Weighting:} The importance of feature weighting has already been discussed in section \ref{fwt}. It is a fraction that lies between 0 and 1, with 0 being the least important and 1 the most important feature. The extent of mutation is multiplied by this fraction to determine how much a specific feature changes by. Generally, we would like to make the largest change to the most crucial feature.
\item \textit{Information Pruning:} In addition to the weighting the features, we explored the possibility of mutating only the most important features. The feature weights are already known to us, we sort the features in descending order of their weights and mutate only the first X\%.
\ei
It is to be noted however that
